---
title: "CPSA Summary"
format: html
project:
  type: default
  output-dir: "CPSA"
editor: source
execute:
  message: false
  warning: false
  echo: false
---

```{r}
#| label: load-data
source("3_model_votechoice_occupation_oesch_5.R")
theme_set(theme_minimal())
```

## Subjective Social Class

```{r}
#| label: distribution-social-class
#| include: true
ces %>% 
  count(election,sub_class2) %>% 
  group_by(election) %>% 
  mutate(Percent=n/sum(n)) %>% 
  filter(!is.na(sub_class2)) %>% 
  ggplot(., aes(y=fct_relevel(sub_class2, "Working Class"), fill=factor(election, levels=c("1984", "2025")), x=Percent))+
  geom_col( position=position_dodge(reverse=T))+
    #geom_linerange(aes(xmin=pct+(1.96*se), xmax=pct+(1.96*se)))+
  scale_fill_manual(values=c("black", "lightgrey"))+
  labs(x="Percent", y="Subjective Social Class", fill="Election", title="Distribution of Subjective Social Class, 1984 and 2025")
```

There is a greater polarization of subjective social class. This fits only partially with what we know about the economy. It fits with the decline of manual manufacturing, but does not fit with increasing income inequality.

## Concordance 

There is a mixed picture of concordance. There are more objective working class inflators than before but also more objective middle class deflators. Might that be related to housing?

```{r}
#| label: concordance

ces %>% count(election, obj_class2,concordance) %>% 
  group_by(election, obj_class2) %>% 
    filter(!is.na(concordance)) %>% 
  mutate(Percent=n/sum(n)) %>% 
  ggplot(., aes(x=Percent, y=fct_relevel(concordance, "Deflator", "Concordant"), fill=election))+  geom_col( position=position_dodge(reverse=T))+labs(x="Percent", y="Objective Social Class", fill="Election", title="Distribution of Subjective Social Class, 1984 and 2025")+facet_grid(~fct_relevel(obj_class2,"Working Class"), scales="free_y")+scale_fill_grey()+theme(legend.position="bottom")
```

- Rates of concordance have gone down for working and middle class people
- Rates of deflation amongst middle class people have gone *up*
- Rates of inflation amongst working and middle class people have also gone *up*

Overall there is a *polarization* of class sensibilities. These changes are statistically significant.  Is this related to housing? 


```{r}
#| label: model-concordance
library(nnet)
# Model only middle class (can deflate, be concordant, or inflate)
mod_middle <- multinom(concordance ~ election, data = subset(ces, obj_class2 == "Middle Class"))

# Model only working class (can only be concordant or inflate)
mod_working <- glm(concordance ~ election, data = subset(ces, obj_class2 == "Working Class"),
                   family = binomial)  # binary: Concordant vs Inflator

# Model only upper class (can only be concordant or deflate)
mod_upper <- glm(concordance ~ election, data = subset(ces, obj_class2 == "Upper Class"),
                 family = binomial)  # binary: Concordant vs Deflator

# Binary models together (working class and upper class)
modelsummary(
  list(
    "Working class\n(ref: Concordant)" = mod_working,
    "Upper class\n(ref: Concordant)"   = mod_upper
  ),
  exponentiate  = TRUE,
  statistic     = "conf.int",
  stars         = TRUE,
  coef_rename   = c("election2025" = "2025 (ref: 1984)"),
  gof_map       = c("nobs", "logLik", "AIC"),
  title         = "Odds of inflation (working class) and deflation (upper class) vs. concordance"
)
```
```{r}
#| label: middle-class-inflation
# Multinomial model (middle class)
modelsummary(
  list("Middle class" = mod_middle),
  shape         = term ~ model + response,
  exponentiate  = TRUE,
  statistic     = "conf.int",
  stars         = TRUE,
  coef_rename   = c("election2025" = "2025 (ref: 1984)"),
  gof_map       = c("nobs", "logLik", "AIC"),
  title         = "Odds of deflation and inflation vs. concordance (middle class)"
)
```

### Concordance and housing.

```{r}
#| label: concorance-own-models-working-upper
mod_own_working_84<-glm(concordance~own_rent, data=subset(ces, obj_class2=="Working Class"&election=="1984"), family="binomial")
mod_own_working_25<-glm(concordance~own_rent, data=subset(ces, obj_class2=="Working Class"&election=="2025"), family="binomial")
mod_own_upper_84<-glm(concordance~own_rent, data=subset(ces, obj_class2=="Upper Class"&election=="1984"), family="binomial")
mod_own_upper_25<-glm(concordance~own_rent, data=subset(ces, obj_class2=="Upper Class"&election=="2025"), family="binomial")
mod_own_middle_84<-multinom(concordance~own_rent, data=subset(ces, obj_class2=="Middle Class"&election=="1984"))
mod_own_middle_25<-multinom(concordance~own_rent, data=subset(ces, obj_class2=="Middle Class"&election=="2025"))


```
```{r}
#| label: print-own-concordance-models-working-upper
mod_own_list<-list("Working Classs\n1984"=mod_own_working_84, "Working Classs\n2025"=mod_own_working_25, "Upper Classs\n1984"=mod_own_upper_84, "Upper Class\n2025"=mod_own_upper_25)
modelsummary(mod_own_list, stars=T, coef_rename = c("own_rentOwn"="Ownership"))
```
```{r}
#| label: print-own-concordance-models-middle
mod_own_middle_list<-list("Middle Class 1984"=mod_own_middle_84, "Middle Class 2025"=mod_own_middle_25)
modelsummary(mod_own_middle_list, stars=T, shape=term~model+response, coef_rename=c("own_rentOwn"="Ownership"))


```
