---
title: "Oesch Coding Scheme and CES25"
format: html
editor: source
---

```{r}
#| label: load-data

# remotes::install_github("sjkiss/cesdata2", force=T)
library(cesdata2)
library(kableExtra)
library(modelsummary)
library(crosstable)
library(nnet)
library(haven)
#data("cesdata2")
data("ces25b")
```
In the module that we purchased we have `r nrow(ces25b)` respondents. Of those, we have pre-coded NOC codes for `r nrow(subset(ces25b, !is.na(NOC21_5)==T))` respondents, so we do have a sample-size issue. 

We will get another 186 because there are 186 open-ended responses to the occupation question. 

Oesch has multiple categorizations, but I have implemented Table 2 of [this](https://people.unil.ch/danieloesch/files/2020/11/Description_SocialClassPackage_ESS_25Nov2020.pdf)  which is an eight category scheme. The only difference that I have done is that I have just had one self-employed category.  

@tbl-oesch shows the distribution of these responses.

```{r}
#| label: tbl-oesch
#| tbl-cap: "Distribution of Oesch occupation classes"

table(ces25b$occupation_oesch)
```


```{r}
#| label: recodes

ces25b$vote<-as_factor(ces25b$vote)
ces25b$vote<-factor(ces25b$vote, levels=c("Conservative", "Liberal", "NDP", "Bloc", "Green", "Other"))


ces25b$occupation_oesch<-factor(ces25b$occupation_oesch, levels=c("Production workers",
                                         "Office clerks",
                                         "Service workers",
                                         "Self-employed",
                                         "Technical Professionals",
                                         "(Associate) Managers",
                                         "Socio-cultural (semi) Professionals"))

#Set Greens to be other
ces25b$vote2<-car::Recode(ces25b$vote, "'Green'=NA;
                     'Other'=NA", levels=c(  "Conservative","Liberal", "NDP", "Bloc"))

```
This is the cross-tab of vote and occupation

```{r}
#| label: tbl-vote-crosstab
#| results: "markup"

crosstable(ces25b, cols=occupation_oesch, by=vote2) %>% kable()

```
If we model vote as a function of this class, then we get this. 


