---
title: "Checking Missing Values"
format: html

editor: visual
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: load-libraries
library(tidyverse)
library(srvyr)
library(haven)
library(labelled)
```

```{r}
#| label: load-data
#| results: hide
library(cesdata2)

ces25b %>% 
  mutate(valid_oesch=!is.na(occupation_oesch)) %>% 
  count(cps25_employment, valid_oesch) %>% as_factor()
```



## CES25

There are `r nrow(ces25b)` respondents in the module. There are three variables to consider.

1.  `occupation_code` is the 5-digit NOC code respondents selected in the menu that we provided. It was offered to all respondents, retired, student or not. This variable is 5-digits long if respondents selected a code. It is stored as a character variable, so if respondents picked nothing, then it is just empty. 
2.  `pes_occ_select_2` is the open-ended text response that was offered to respondents at the end of the menu selection process if they could not see their response. This variable has the response `-99` if it was seen but not selected, sometimes it is blank, and many occasions a text response such as `retired` or `yes` or actual occupations. We coded those occupations as best we could.
3.  `NOC21_5`. This variable is the combination either of `occupation_code` from the survey menu or the coding process by the RAs.

I'm going to count how many valid values there are on each variable by employment status.

```{r}
#| label: noc-occupation-code 
library(crosstable)

ces25b %>% 
  mutate(occupation_code_valid=case_when(
    nchar(occupation_code)==5~"Valid 5 digit NOC code from menu",
    nchar(occupation_code)!=5~"No Valid 5 digit NOC code from menu"
  ))->ces25b
crosstable(ces25b, occupation_code_valid) %>% as_flextable(., keep_id=F)

```

Unfortunately, only 43% of respondents successfully used the menu.

Now, I'm going to check the open-ended responses in `pes25_occ_select_2`. I'm going to create another variable that codes those responses into three categories:

1.  No response
2.  Saw menu but no open response
3.  Some open response.

```{r}
#| label: select-2-summary
#| results: 'markup'
ces25b %>% 
  mutate(occupation_open=case_when(
    # If pes25_occ_select2 has no characters then there is no response
    nchar(pes25_occ_select_2)==0~"No response",
    #If they 
   pes25_occ_select_2=="-99"~"Saw menu but no open response",
   TRUE~"Some open response"
  )) ->ces25b
crosstable(ces25b,occupation_open)%>% as_flextable(., keep_id=F)
```

We can crosstab these variables and this will tell us what the relationship between `occupation_code` and `pes_occ_select_2` is.

```{r}
#| label: crosstab
crosstable(ces25b,occupation_open~occupation_code_valid)%>% as_flextable(., keep_id=F)

```

Who are those 350 respondents who have a valid 5 digit code from the menu and provided some open response?

```{r}
#| label: check-valid-5-digit-codes
ces25b %>% 
  filter(occupation_code_valid=="Valid 5 digit NOC code from menu" &occupation_open=="Some open response") %>% 
  select(occupation_code, pes25_occ_select_2, NOC21_5) 
```

To me these look like people who picked a menu response (so there is a valid 5 digit NOC code from the menu) and then validated that response in some way. But they clearly do have a good NOC code.

So, what of the of the 489 ones who do *not* have a valid 5-digit NOC code from the Qualtrics menu, but gave an open-ended response, what happened to them?

```{r}
#| label: check-open-ended-responses

  ces25b %>% 
  filter(occupation_open=="Some open response"&occupation_code_valid=="No Valid 5 digit NOC code from menu") %>% 
  select(occupation_code,pes25_occ_select_2, NOC21_5)
```

These are the respondents that we coded. They provided open-ended job titles and we manually coded them into NOC codes.

Now, each of these NOC codes were slotted into Oesch codes. I'm pretty sure we were able to capture all of those. 

But there are a mix of employed, retired and unemployed people who provided NOC codes. What is the distribution of Oesch codes across those employment statuses.
I have currently eliminated all employment status restrictions in the code to make the `occupation_oesch` variable. All of these people somehow gave some kind of valid NOC21 code that can be mapped to a class category. I think this is as big as we can go. 

The distribution of valid oesch codes by employment status is as follows:

```{r}
#| label: count-valid-noc-codes-by-emp-status

ces25b %>% 
  mutate(valid_oesch=!is.na(occupation_oesch), valid_noc=!is.na(NOC21_5))->ces25b
crosstable(ces25b, as_factor(cps25_employment)~valid_oesch)%>% as_flextable(., keep_id=F)

```


That will give us 296 retired people, 42, unemployed, 26 students, 14 homemakers and 28 disabled people. That's `r 296+42+26+14+28`.

For a grand total, we will have:

```{r}
#| label: sample-size-2025

table(!is.na(ces25b$occupation_oesch))

```
If we compare this to 1984 we can see what happens. 

```{r}
#| label: check-1984

ces84 %>% 
count(VAR524, valid_oesch=!is.na(occupation_oesch)) %>% as_factor()
```
One thing to note, there is a huge difference in having a phone interviewer soliciting occupation codes compared to the web mode. We got around 1/4 of the respondents who worked providing a valid occupation code. Having an interviewer probing a response makes a huge difference. Also the 802 respondents who responded `housewife` do not have valid oesch codes because they truly do not have occupation codes. They were assigned a `31` in the occupation variable. Look:

```{r}
#| label: show-housewives
ces84 %>% 
  filter(SOC==31) %>% 
  select(VAR524, SOC) %>% as_factor() %>% 
  slice_sample(n=25)
```

The retired people however were offered the chance to provide an occupation.

```{r}
#| label: show-retired

ces84 %>% 
  filter(VAR524==3) %>% 
  select(VAR524, SOC, occupation_oesch) %>% as_factor() %>% 
  slice_sample(n=25)

```


But, we can use the same logic as in ces25b and provide an expansive definition of who class as anyone who provided a valid occupation. That will give us this sample size:

```{r}
#| label: sample-size1984
table(!is.na(ces84$occupation_oesch))
```
A lot lower, but more than where we started with!
